---
title: "Graph Analysis Project"
author: 
- "NASR Rodrigue"
- "NENOVSKY Nikola"
date: "2024-03-28"
output: 
  pdf_document:
    toc: yes
  html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Research question: Study of a network formation mechanism

*In this project, you will investigate a dyadic mechanism driving the formation of a real network. This mechanism can be related to assortativity or dissortativity on a node attribute (similar or dissimilar individuals tend to be connected), to the sociability or popularity of a node attribute (some individuals connect to more people or receive more connections), or to another network variable (individuals connected through another network tend to be connected). Please leave the instructions in italic. Attach your code as a RmD file.*

## Data importation and first plot

```{r}
library(tidyverse)
library(igraph)
library(igraphdata)
library(RColorBrewer)
library(dplyr)
library(ggplot2)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
